---
title: "Brauchle_HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ISLR)
weekly <- Weekly
```

```{R} 
a_model <- glm(Direction ~ Lag1 + Lag2, data = Weekly, family = "binomial")
summary(a_model)

```
(a) The model including the full dataset only had Lag2 as a significant predictor of Direction (z = 2.27, p < 0.05). 

```{r}
weekly_2 <- Weekly[-1,]
b_model <- glm(Direction ~ Lag1 + Lag2, data = weekly_2, family = "binomial")
summary(b_model)
```
(b) Removing the first observation does not change the model very significantly, and Lag2 remains as the only significant predictor of Direction (z = 2.29, p < 0.05). 

```{r}
#c
glm.probs = predict(b_model, weekly[1,], type="response") 
if (glm.probs > 0.5) {
  pred_dir = "Up"
}else {
  pred_dir = "Down"
}
pred_dir
pred_dir == Weekly[1,]$Direction

```
(c) The first observation was predicted as "Up." However, this observation was not classified correctly, as the true direction of the first observation is "Down."

```{r}
#d 
errors = vector()
for (i in (1:nrow(weekly))) {
  loop_model <- glm(Direction ~ Lag1 + Lag2, data = Weekly[-i,], family = "binomial")
  glm.probs = predict(loop_model, weekly[i,], type="response")
  
  if (glm.probs > 0.5) {
    pred_dir = "Up"
  } else {
    pred_dir = "Down"
  }

  if (pred_dir != Weekly[i,]$Direction){
    errors[i] = 1
  } else {
    errors[i] = 0
  }
}
errors
```


```{r}
mean(errors)
```

(e) Using the leave-one-out-cross-validation method produced an error estimate of 45%, so only about 55% of observations were classified correctly. 
